---
title: "Course Project 1 - Reproducible Research"
author: "Arushi Nigam"
---
  
###Assignment Instructions
1.Code for reading in the dataset and/or processing the data
2.Histogram of the total number of steps taken each day
3.Mean and median number of steps taken each day
4.Time series plot of the average number of steps taken
5.The 5-minute interval that, on average, contains the maximum number of steps
6.Code to describe and show a strategy for imputing missing data
7.Histogram of the total number of steps taken each day after missing values are imputed
8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
9.All of the R code needed to reproduce the results (numbers, plots, etc.) in the report


##Step 1
##Code for reading in the dataset and/or processing the data
```{r, echo = TRUE}
activity<-read.csv("activity.csv")
```


##Step 2
##Histogram of the total number of steps taken each day
```{r, echo = TRUE}
library(dplyr)
library(ggplot2)
png("plot1.png", width = 480, height = 480)
activityData <- read.csv("activity.csv")
relData <- activityData %>% group_by(date) %>% summarize_all(sum, na.rm = TRUE)
hist(relData$steps, xlab = "Total number of steps per day")
dev.off()

```


##Step 3
##Mean and median number of steps taken each day

```{r, echo = TRUE}
averageSteps<- mean(relData$steps)
medianSteps <- median(relData$steps)
print(averageSteps)
print(medianSteps)
```

##Step 4
##Time series plot of the average number of steps taken
```{r, echo = TRUE}
png("plot2.png", width = 480, height = 480)
library(dplyr)
library(ggplot2)
activityData <- read.csv("activity.csv")
relData <- activityData %>% group_by(interval) %>% summarize_at(vars(steps), mean, na.rm = TRUE)
g <- ggplot(data = relData, aes(interval, steps)) + geom_line()
g
dev.off()
```


##Step 5
##The 5-minute interval that, on average, contains the maximum number of steps

```{r, echo = TRUE}
interval <- relData$interval[which.max(relData$steps)]
print(interval)
```


##Step 6
Code to describe and show a strategy for imputing missing data
There are multiple strategies to deal with multiple value imputations.
I will be using the mean of 5 minute intervals as a substitute for the respective NA values.
I'm also displaying the respective mean and median values.
```{r, echo = TRUE}
png("plot3.png", width = 480, height = 480)
library(dplyr)
library(ggplot2)
library(zoo)
activityData <- read.csv("activity.csv")
missingValues <- sum(is.na(activityData$steps))
print(missingValues)
activityData$steps <- na.aggregate(activityData$steps, activityData$interval)
relData <- activityData %>% group_by(date) %>% summarize_all(sum, na.rm = TRUE)
hist(relData$steps, xlab = "Total number of steps per day")
dev.off()
averageSteps <- mean(relData$steps)
medianSteps <- median(relData$steps)
print(averageSteps)
print(medianSteps)
Q6.6 <- activityData


```

By this point, from the plot, that the missing values have a very disctinct pattern. For every interval, there are consistantly 8 missing values. For the date, there are consistantly 288 missing values. And in total, there are 8 dates that have missing value. 
In total, there are 2304 missing values as shown.

## Step 7
Histogram of the total number of steps taken each day after missing values are imputed

```{r, echo = TRUE}
png("plot4.png")
qplot(Q6.6$steps,geom="histogram",main="Total steps taken histogram post imputation",xlab="Steps",ylab="Count")
dev.off()
Q7 <- Q6.6

```




## Step 8
Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r, echo = TRUE}
png("plot5.png", width = 480, height = 480)

activityData <- read.csv("activity.csv")
activityData$steps <- na.aggregate(activityData$steps, activityData$interval)

activityData$date <- as.Date(activityData$date)

activityData <- mutate(activityData, day = weekdays(activityData$date))
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
activityData <- mutate(activityData, day_type = factor((activityData$day %in% weekdays), levels = c(FALSE, TRUE), labels = c("Weekend", "Weekday")))


activityData <- activityData %>% group_by(interval, day_type) %>% summarize_at(vars(steps), mean, na.rm = TRUE)

g <- ggplot(activityData, aes(x = interval, y = steps))
g <- g + facet_grid(day_type ~ .) + geom_line() + ggtitle("Mean steps per interval on Weekdays and Weekends")
g

dev.off()
```
